# Data access {#sec-data-access}

This section is intended as a "REDCap data manipulation 101" for pre-graduate students in the group and as an intro to getting RStudio set up for work with REDCap data.

In the following is a guide to get started working on REDCap data in R using the [REDCapR-package](https://ouhscbbmc.github.io/REDCapR/).[@R-REDCapR]

Following the package link provides an extensive description on workflows and functionality. However, sometimes less is more, and so goes.

A more thorough discussion on REDCap data can be found later as well as a chapter on tips to building a REDCap database.

## Getting access {#sec-getting-access}

Before you are able to get any data out of REDCap, make sure that you are able to get in. After you are logged in, make sure, that you have permissions to get an API-key. The API-key is a alpha-numeric key used to verify yourself to the REDCap server, when requesting data export.

The API key equals user name and password, and circumvents the two-factor verification process. Naturally, the API key should be kept in a safe way.

::: callout-warning
To be clear: **NEVER STORE THE API-KEY AS PLAIN TEXT.**
:::

The `library(keyring)`[@R-keyring] offers a safe and easy way to store API keys and other secrets. Please copy and run the code below:

```{r api-storage, eval=FALSE}
require(keyring)
key_set("API_KEY", prompt = "API key: ")
```

```{r uri-storage, eval=FALSE}
key_set("DB_URI", prompt = "Data base uri: ")
```

Do note, that the name "API_KEY" can be changed to whatever you prefer. Note that the function will overwrite any existing keys with the same name without any warning (!).

##  {#sec-downloading-data}
